---
title: "Gender Wage Inequality in STEM"
author: "Lydia Gibson, Sara Hatter & Ken Vu"
date: 'April 28, 2022'
output:
  beamer_presentation: default
  ioslides_presentation: default
---

# Introduction

```{r echo=FALSE}
dat1 <- read.csv("women-stem.csv")
library(pacman)
suppressWarnings(p_load(dplyr, ggplot2, ggpubr, scales, MASS, car, lmtest))
options(scipen = 100) # remove scientific notation
```

```{r echo=FALSE}
dat2 <- dat1[,-c(1,2,3)] # remove Major_code, Rank
#head(dat2)
#colnames(dat2)
```

```{r echo=FALSE}
#set major category as a factor
dat2$Major_category <- as.factor(dat2$Major_category)
#for (i in colnames(dat2[,c(1,4:8)])){
#  dat2[[i]] <- as.numeric(dat2[[i]])
#}
```


## Research Question: 

What associations exist within STEM college majors that have an effect on median wages?

## Goals: 

* To explore the data for STEM college majors.
* To create a predictive model for median wage.

# Data Description

The data was obtained from the American Community Survey 2010-2012 Public Use Microdata Series and has been already subsetted to only concern STEM majors (particularly with an interest in women majoring in STEM).  For each row in the data set (which represents one major), there’s a collection of details and statistics about the major, such as the type of major (i.e. Engineering, Health Science, etc), the proportion of women in the sample of individuals working in that particular field, and other relevant pieces of information.

# Data set

The dimensions of the data set are 76 rows by 9 columns.

```{r echo=FALSE}
head(dat1, n=3)

```
* Link to data set: https://github.com/fivethirtyeight/data/blob/master/college-majors/women-stem.csv

# Last 3 rows of data

```{r echo=FALSE}
tail(dat1, n=3)
```

* Link to data set: https://github.com/fivethirtyeight/data/blob/master/college-majors/women-stem.csv

# Variables

We chose `Median`: Median earnings of full-time, year-round workers as our response variable and our predictor variables are:

- `Rank`: Rank by median earnings

- `Major_code`: Major code, FO1DP in ACS PUMS

- `Major`: Major description

- `Major_category`: Category of major from Carnevale et al

- `Total`: Total number of people with major

- `Men`: Male graduates

- `Women`:Female graduates

- `ShareWomen`: Women as share of total

# Methods and Results: EDA

Within the STEM majors, median wage ranged from $\$26,000$ for Zoology and $\$110,000$ for Petroleum Engineering. 

```{r echo=FALSE}
summary(dat2$Median)
```


Each individual major within the data set is further categorized into one of five major categories:

* Biology & Life Sciences

* Computers & Mathematics

* Engineering

* Health

* Physical Sciences


# Median Wage by Major Category

```{r echo=FALSE}
boxplot(Median~Major_category,data=dat2)
```

# Test differences between major categories

Based on our boxplot, we noticed there may be a significant difference between median wage by major category so we ran an ANOVA to test our hypothesis.

```{r}
# anova of major categories
```


# Share Women by Major Category

During our exploratory data analysis we noticed an interesting association between share of women within a given major category.

```{r echo=FALSE, warning=FALSE, message=FALSE}

# Get totals for men and women for each major category
dat_stats <- rbind(
  
  # Get totals for men
  dat2 %>% group_by(Major_category) %>% 
  summarize(Grand_Total = sum(Men), Proportion=Grand_Total/sum(Total)) %>% 
    mutate(Sex="Men"),
  
  # Get totals for women
  dat2 %>% group_by(Major_category) %>% 
  summarize(Grand_Total = sum(Women), Proportion=Grand_Total/sum(Total)) %>% 
    mutate(Sex="Women")
)
#dat_stats

dat_stats %>% ggplot(aes(x=Major_category,y=Grand_Total,fill=Sex)) +
  stat_summary(geom = "bar", position="dodge") +  # stack side by side bars 
  # Get labels for bars
  geom_text(aes(label=percent(Proportion)), position = position_dodge(width=.9),
            vjust=-1, size=3) + 
  theme(axis.text.x=element_text(angle = 7.5),  # get x axes labels to fit
        plot.title=element_text(size=17, hjust=0.5)) + # center title
  scale_y_continuous(limits=c(0, 4.25*10^(5))) + # increase size to fit labels
  labs(x="Major Category", y="Number of People",
       title="Major Category Proportions by Sex")
```

# Checking Assumptions

Before beginning our analysis, we began by exploring the normality within our response variable, `Median`. We notices that there was some skewing, so we decided to do a log transformation.

```{r}
#test normality of median

```


# Scaterplot Matrix

```{r echo=FALSE}
pairs(log(Median)~., data=dat2[,-c(1)])
```

# Remove variables from Model

As noticed in our scatterplot matrix, there may be an issues with multicollinearity between `Total`, `Men`, `Women` and `ShareWomen`, so we ran some analysis to see which of these predictors could be removed from our model.

# Conclusion

In conclusion

- Point 1

- Point 1

- Point 1


# Code Appendix

For supplementary R script, visit 

- https://github.com/lgibson7/Gender-Wage-Inequality-in-STEM